Survival Analysis

### 2014-04-28 paper.Rnw
Added visualization of bfeed survival

\begin{document}
\SweaveOpts{concordance=TRUE}
\maketitle
<<echo = FALSE>>=
library(ggplot2)
library(gridExtra)
library(survival)
library(KMsurv)
library(xtable)

@@

data(bfeed)
@

@@

Breastfeeding duration is measured in weeks. The study tracks the following factors: race (white, black, other), poverty, somking, drinks alcohol, age of mother, mother's year of birth, mother's education level, and if the mother has prenatal car after the 3rd month of pregnancy.

@@

@@

\section*{Results}
% Exploratory Analysis (“graphical”: graphs, descriptive statistics)
% * Survival of some college vs no college
% * Survival by poverty (before 27 weeks, after 27 weeks)
% Statistical Analysis (“analytical”: hypotheses, test statistics, p-values, conclusions)
% Validation of Assumptions (graphs and/or tests)

>>>>

\begin{document}
\SweaveOpts{concordance=TRUE}
\maketitle
<<echo = FALSE>>=
library(ggplot2)
library(gridExtra)
library(survival)
library(GGally)
library(KMsurv)
library(xtable)

@@

data(bfeed)

bfeed.short <- bfeed
bfeed.short$duration <- with(bfeed.short, ifelse(duration > 36, 36, duration))
@

@@

Breastfeeding duration is measured in weeks. An event (delta) is defined as full weaning of the infant from breast milk. The study tracks the following factors: race (white, black, other), poverty, somking, drinks alcohol, age of mother, mother's year of birth, mother's education level, and if the mother has prenatal car after the 3rd month of pregnancy.

@@

The range of durations for breastfeeding is from 1 week to 192 weeks. Only 12\% of subjects had breastfeeding durations greater than 36 weeks. If we were to graph $\hat S(t)$, 20\% of the x-axis will contain 88\% of the data, which makes detecting patterns in the data visually very difficult. For this reason, durations greater than 36 weeks will be treated as 36 weeks.

@@

\section*{Results}

Below is a graph of survival data for all points with no explanatory variables in the model.

<<bfeed-survival, echo = FALSE, fig = TRUE, include = FALSE>>=
bfeed.fit <- survfit(Surv(duration, delta) ~ 1, data = bfeed.short)

ggsurv(bfeed.fit) +
  ggtitle('Breast Feeding Duration') +
  xlab('Duration in Weeks') +
  ylab('Survival') +
  geom_vline(aes(xintercept = 27),
    colour = 'red',
    linetype = 'longdash')
@
\begin{figure}[h]
\begin{center}
\includegraphics[keepaspectratio=true, width=3in]{paper-bfeed-survival.pdf}
\end{center}
\caption{Breast Feeding Survival Data}
\label{fig:bfeed-survival}
\end{figure}

The red dashed line represents the Food and Drug Administration's recommended duration of breastfeeding, 6 months. The steep drop off at 36 weeks is due to the data clipping used to make the graphs easier to interpret. The red crosses indicate censored observations and the dashed black line is the confidence band of the Kaplan-Meier estimator.

% Exploratory Analysis (“graphical”: graphs, descriptive statistics)
% * Survival of some college vs no college
% * Survival by poverty (before 27 weeks, after 27 weeks)
% Statistical Analysis (“analytical”: hypotheses, test statistics, p-values, conclusions)
% Validation of Assumptions (graphs and/or tests)
